---
title: "Data 622 - Homework 2"
author: "Leticia Salazar"
date: "October 29, 2023"
output: 
  html_document:
    theme: sandstone
    highlight: haddock
  pdf_document:
    latex_engine: xelatex
    dev: cairo_pdf
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
$~$

#### Pre-work
1. Read this blog: https://decizone.com/blog/the-good-the-bad-the-ugly-of-using-decision-trees which shows some of the issues with decision trees. 
2. Choose a dataset from a source in Assignment #1, or another dataset of your choice.


#### Assignment work
1. Based on the latest topics presented, choose a dataset of your choice and create a Decision Tree where you can solve a classification problem and predict the outcome of a particular feature or detail of the data used.
2. Switch variables* to generate 2 decision trees and compare the results. Create a random forest for regression and analyze the results.
3. Based on real cases where desicion trees went wrong, and 'the bad & ugly' aspects of decision trees (https://decizone.com/blog/the-good-the-bad-the-ugly-of-using-decision-trees), how can you change this perception when using the decision tree you created to solve a real problem?


#### Deliverable
1. Essay (minimum 500 word document)
Write a short essay explaining your analysis, and how you would address the concerns in the blog (listed in pre-work)
2. Exploratory Analysis using R or Python (submit code + errors + analysis as notebook or copy/paste to document)

Due Date: Sunday October 29th, end of day
* We are trying to just train 2 different decision trees - so swicth the feature used for the first node (split) to force a different decision tree.

$~$

#### Load Libraries:
Below are the libraries used to complete this assignment
```{r}

```

$~$

#### Load Data:
The data chosen from Excel BI Analytics were the 100 sales records for the small and 5000 sales records for the large. The data sets are included in my [GitHub](https://github.com/letisalba/Data-622/tree/master/Homework-1/csv) and read into R.
```{r}

```

$~$

### The Data:


$~$

### Data Exploration:


$~$
### Data Preparation:

$~$

### Model Building:


$~$

### Conclusion: